---
title: "reading_data_from_web"
output: html_document
date: "2023-10-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## AirBnB data brainstorming

### Brainstorm questions

- Where are AirBNBs expensive? 
  - Borough, Neighborhood? 
  - do other factors (room type) affect price? what about rating?
- how long are AirBNBs active? 
- which units have the most availability? 
- how many apts are run by one host? 
- where are AirBNBs? 
  - Borough? Neighborhood? 
  - Lat and Long? 
  
```{r}
library(tidyverse)
library(p8105.datasets)

data(nyc_airbnb)

nyc_airbnb = 
  nyc_airbnb |> 
  rename(borough = neighbourhood_group) |> 
  mutate(stars = review_scores_location / 2)
```

## Some EDA
Where are AirBNBs? Borough, neighborhood? lat and long? 

```{r}
nyc_airbnb |> 
  group_by(borough, neighbourhood) |> 
  count(borough, neighbourhood) |> 
  arrange(desc(n))

nyc_airbnb |> 
  count(borough)
```

where are the most expensive AirBnBs by borough? 

```{r}
nyc_airbnb |> 
  group_by(borough, room_type) |> 
  summarize(median_price = median(price)) |> 
  pivot_wider(
    names_from = room_type,
    values_from = median_price
  )
```

some visualizations to go with this

```{r}
nyc_airbnb |> 
  ggplot(aes(x = price)) + 
  geom_histogram() +
  facet_grid(borough ~ room_type)
```

```{r}
nyc_airbnb |> 
  filter(price >= 9500)
```

```{r}
nyc_airbnb |> 
  filter(price < 1000, room_type == "Entire home/apt") |> 
  ggplot(aes(x = price)) + 
  geom_histogram() +
  facet_grid(borough ~ room_type)
```

where in manhattan are airbnbs most expensive? 

```{r}
nyc_airbnb |> 
  filter(
    price < 1000,
    room_type == "Entire home/apt",
    borough == "Manhattan") |> 
  group_by(neighbourhood) |> 
  summarize(
    n_apt = n(),
    mean_price = mean(price)) |> 
    arrange(desc(mean_price))
``` 





