---
title: "Strings and Factors"
output: html_document
date: "2023-10-10"
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Strings and manipulations

```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "jeff")
```

^ `str_detect`, you want to use to find something that exists. use in conjunction with `filter` to get rid of things that don't have a specific pattern match 

```{r}
str_replace(string_vec, "jeff", "Jeff")
```

^ `str_replace` replaces characters in string vectors

```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "i think")
str_detect(string_vec, "^i think")
str_detect(string_vec, "i think$")

```

^ use `^` to indicate the line starts there, to indicate the first three rather than in the entire sentence. `$` does the opposite, for the end of the sentence

```{r}
string_vec = c(
  "Time for a Pumpkin Spice Latte!",
  "went to the #pumpkinpatch last weekend",
  "Pumpkin Pie is obviously the best pie",
  "SMASHING PUMPKINS -- LIVE IN CONCERT!!"
  )

str_detect(string_vec,"Pumpkin")
str_detect(string_vec,"pumpkin")
str_detect(string_vec,"[Pp]umpkin")
```

^`[]` indicates that you can match one thing or the other thing (both? only one or the other?)

```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "[0-9]")
str_detect(string_vec, "^[0-9]")
str_detect(string_vec, "^[0-9][a-zA-Z]")
```

^`[]` any instance that matches any number 0-9, add `^` to indicate first line, can layer `[]` to build

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")
```

^ `.` selects anything from the selected choices (eg "7, 11")

```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")
str_detect(string_vec, "\\[[0-9]")
```

^ in order to detect "[", it tyipcally gets interpreted as a regular expression function. "[" is a special character and needs to be designated by "\\". ("\" is also a special character, so add another one)














